---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

#worldfootballR

## Overview

This package is designed to allow users to extract various world football results and player statistics data from fbref.com


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR", ref = "main")
```

```{r load_libs}
library(worldfootballR)
```


## Usage

The functions available in this package are designed to enable the extraction of world football data.

### Get match results

To get the match results (and additional metadata) for all games for a tier-1 league season, the following function can be used:

```{r get_match_results}
# function to extract chess.com game data
serieA_2020 <- get_match_results(country = "ITA", gender = "M", season_end_year = "2020")
head(serieA_2020, 10)
```


### Get Season Team Stats

The `get_season_team_stats` function allows the user to return a data frame of different stat types for all teams in a tier-1 league season.

Note, some stats may not be available for all leagues. The big five European leagues should have all of these stats.

The following stat types can be selected:

* *league_table*
* *league_table_home_away*
* *standard*
* *keeper*
* *keeper_adv*
* *shooting*
* *passing*
* *goal_shot_creation*
* *defense*
* *possession*
* *playing_time*
* *misc*

```{r get_season_team_stats}
# function to extract season teams stats
prem_2020_shooting <- get_season_team_stats(country = "ENG", gender = "M", season_end_year = "2020", stat_type = "shooting")
prem_2020_shooting
```
