---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

#worldfootballR

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/JaseZiv/worldfootballR.svg?branch=main)](https://travis-ci.org/JaseZiv/worldfootballR)
<!-- badges: end -->

## Overview

This package is designed to allow users to extract various world football results and player statistics data from fbref.com


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
library(tidyverse)
```


## Usage

The functions available in this package are designed to enable the extraction of world football data.

There are three main categories of data extract functions in this package:

* Match-level statistics (team and player)
* Season-level statistics (team and player)
* League / Team metadata

### Match-level statistics

#### Get match results

To get the match results (and additional metadata) for all games for a tier-1 league season, the following function can be used:

```{r get_match_results}
# function to extract chess.com game data
serieA_2020 <- get_match_results(country = "ITA", gender = "M", season_end_year = 2020)
glimpse(serieA_2020)
```

**More than one league season**

The `get_match_results` function can be used to get data for multiple seasons/leages/genders/etc also:

```{r get_match_results_multiple, eval=FALSE}
big_5_2020_results <- get_match_results(country = c("ENG", "ESP", "ITA", "GER", "FRA"),
                                        gender = "M", season_end_year = 2020)
```


#### Get match report

This function will return similar results to that of `get_match_results()`, however `get_match_report()` will provide some additional information. It will also only provide it for a single match, not the whole season:  

```{r get_match_report}
# function to extract chess.com game data
liv_mci_2020 <- get_match_report(match_url = "https://fbref.com/en/matches/47880eb7/Liverpool-Manchester-City-November-10-2019-Premier-League")
glimpse(liv_mci_2020)
```


#### Get match lineups

This function will return a dataframe of all players listed for that match, including whether they started on the pitch, or on the bench.  

```{r get_match_lineups}
# function to extract chess.com game data
liv_mci_2020_lineups <- get_match_lineups(match_url = "https://fbref.com/en/matches/47880eb7/Liverpool-Manchester-City-November-10-2019-Premier-League")
glimpse(liv_mci_2020_lineups)
```

***

### Season-level statistics


#### Get Season Team Stats

The `get_season_team_stats` function allows the user to return a data frame of different stat types for all teams in tier-1 league seasons.

Note, some stats may not be available for all leagues. The big five European leagues should have all of these stats.

The following stat types can be selected:

* *league_table*
* *league_table_home_away*
* *standard*
* *keeper*
* *keeper_adv*
* *shooting*
* *passing*
* *goal_shot_creation*
* *defense*
* *possession*
* *playing_time*
* *misc*

```{r get_season_team_stats}
# function to extract season teams stats
prem_2020_shooting <- get_season_team_stats(country = "ENG", gender = "M", season_end_year = "2020", stat_type = "shooting")
glimpse(prem_2020_shooting)
```


**More than one league season**

The `get_season_team_stats` function can be used to get data for multiple seasons/leages/genders/etc.

Important to note, this function can only be used for one `stat-type` at a time, however all other parameters can have multiple values:

```{r get_season_team_stats_multiple, eval=FALSE}
big_5_2020_possessions <- get_season_team_stats(country = c("ENG", "ESP", "ITA", "GER", "FRA"),
                                        gender = "M", season_end_year = 2020, stat_type = "possession")
```
