---
title: "Extracting season data"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting season data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics data from fbref.com and valuations and transfer data from transfermarkt.com.


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
library(dplyr)
```

***


## Usage

The functions available in this package are designed to enable the extraction of world football data.

There are four main categories of data extract functions in this package:

* fbref helper functions to extract key URLs
* Match-level statistics (team and player)
* Season-level statistics (team and player)
* Player market vluations and transfer histories

This vignette will cover season-level data, typically totals for all games played within seasons.

***

## Team Stats

This section will cover the functions to aid in the extraction of season team statistics.

### Get Season Team Stats

The `get_season_team_stats` function allows the user to return a data frame of different stat types for all teams in Domestic leagues [here](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/all_leages_and_cups/all_competitions.csv).

Note, some stats may not be available for all leagues. The big five European leagues should have all of these stats.

The following stat types can be selected:

* *league_table*
* *league_table_home_away*
* *standard*
* *keeper*
* *keeper_adv*
* *shooting*
* *passing*
* *passing_types*
* *goal_shot_creation*
* *defense*
* *possession*
* *playing_time*
* *misc*

```{r get_season_team_stats, eval=FALSE}
# function to extract season teams stats
prem_2020_shooting <- get_season_team_stats(country = "ENG", gender = "M", season_end_year = "2020", tier = "1st", stat_type = "shooting")
# to get shooting stats for the English Championship:
championship_2020_shooting <- get_season_team_stats(country = "ENG", gender = "M", season_end_year = "2020", tier = "2nd", stat_type = "shooting")
```


**More than one league season**

The `get_season_team_stats` function can be used to get data for multiple seasons/leages/genders/etc.

Important to note, this function can only be used for one `stat-type` at a time, however all other parameters can have multiple values:

```{r get_season_team_stats_multiple, eval=FALSE}
big_5_2020_possessions <- get_season_team_stats(country = c("ENG", "ESP", "ITA", "GER", "FRA"),
                                        gender = "M", season_end_year = 2020, tier = "1st", stat_type = "possession")
```

***

## Player Stats

This section will cover the functions available to aid in the extraction of player season data

### Get Player Scouting Report

The `fb_player_scouting_report()` function takes in a player_url as the input, and returns scouting reports for the player selected. The scouting reports are a new feature on fbref, thus only available for the big five Euro leagues:

```{r get_player_scouting_report}
fb_player_scouting_report("https://fbref.com/en/players/3bb7b8b4/Ederson")
```



### Get Player Season Stats

The `fb_player_season_stats()` function allows for the extraction of season totals for selected player URLs and stat_type.

The stat_types available for use in this function are below:

* *standard*
* *shooting*
* *passing* 
* *passing_types*
* *gca*
* *defense*
* *possession*
* *playing_time*
* *misc*
* *keeper*
* *keeper_adv*

```{r player_season_stats, eval=FALSE}
mo_shooting <- fb_player_season_stats("https://fbref.com/en/players/e342ad68/Mohamed-Salah", stat_type = 'shooting')
```


