---
title: "Extracting data from Transfermarkt"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting data from Transfermarkt}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics from the following popular football (soccer) data sites:

* [FBref](https://fbref.com/en/)
* [Transfermarkt](https://www.transfermarkt.com/)
* [Understat](https://understat.com/)


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
```

***

## Usage

Package vignettes have been built to help you get started with the package.

* For functions to extract data from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-fbref-data.html)
* For functions to extract data from Transfermarkt, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-transfermarkt-data.html)
* For functions to extract data from Understat, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-understat-data.html)
* For functions to extract data for international matches from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/fbref-data-internationals.html)

This vignette will cover the functions to extract data from transfermarkt.com


***

## Join FBref and Transfermarkt data

To be able to join data player between FBref and Transfermarkt, `player_dictionary_mapping()` has been created. There are over 6,100 players who have been listed for teams in the Big 5 Euro leagues on FBref since the start of the 2017-18 seasons, with all of these mapped together. This is expected to be updated and grow over time. The raw data is stored [here](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/fbref-tm-player-mapping/output/fbref_to_tm_mapping.csv)

```{r mapped_players}
mapped_players <- player_dictionary_mapping()
dplyr::glimpse(mapped_players)
```


## Transfermarkt Helper Functions

The following section will outline the various functions available to find different URLs to be able to pass through the Transfermarkt suite of functions outlined in this vignette.

### Team URLs

To get a list of URLs for each team in a particular season from transfermarkt.com, the `tm_league_team_urls()` function can be used. If the country/countries aren't available in the main data set, the function can also accept a League URL from transfermarkt.com. To get the league URL, use the filtering options towards the top of transfermarkt.com, select the country and league you want to collect data from, head to that page, and copy the URL.

```{r tm_team_urls, eval=FALSE}
team_urls <- tm_league_team_urls(country_name = "England", start_year = 2020)
# if it's not a league in the stored leagues data in worldfootballR_data repo:
league_one_teams <- tm_league_team_urls(start_year = 2020, league_url = "https://www.transfermarkt.com/league-one/startseite/wettbewerb/GB3")
```


### Player URLs

To get a list of player URLs for a particular team in transfermarkt.com, the `tm_team_player_urls()` function can be used.

```{r tm_player_urls, eval=FALSE}
tm_team_player_urls(team_url = "https://www.transfermarkt.com/fc-burnley/startseite/verein/1132/saison_id/2020")
```


### Staff URLs

To get a list of staff URLs for a particular team(s) and staff role in transfermarkt.com, the `tm_league_staff_urls()` function can be used.

The staff roles that can be passed to the function via the `staff_role` argument are below:

* "Manager" (this will also return caretaker managers)
* "Assistant Manager"
* "Goalkeeping Coach"
* "Fitness Coach"
* "Conditioning Coach"

```{r, tm_staff_urls, eval=FALSE}
# get a list of team URLs for the EPL 2021/22 season
epl_teams <- tm_league_team_urls(country_name = "England", start_year = 2021)
# get all EPL managers for the 2021/22 season
epl_managers <- tm_team_staff_urls(team_urls = epl_teams, staff_role = "Manager")

# get all EPL goal keeping coaches for the 2021/22 season
epl_gk_coaches <- tm_team_staff_urls(team_urls = epl_teams, staff_role = "Goalkeeping Coach")
```


***

## League Season-Level Data

This section will cover the functions to aid in the extraction of season team statistics.


### League Table by Matchdays

To be able to extract league tables for select matchday(s), the below function can be used. 

The function can accept either the country name, season start year and matchday number(s), or for leagues not contained in the [worldfootballR_data](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/transfermarkt_leagues/main_comp_seasons.csv) repository, it can accept the league URL, season start year and matchday number(s).


```{r matchday_league_table, eval=FALSE}
# to get the EPL table after matchday 1 of the 20/21 season:
epl_matchday_1_table <- tm_matchday_table(country_name="England", start_year="2020", matchday=1)

# to get the EPL table after each matchdays from matchday 1 to matchday 35 of the 20/21 season:
epl_matchday_1to35_table <- tm_matchday_table(country_name="England", start_year="2020", matchday=c(1:35))

# to get the EPL table after each matchdays from matchday 1 to matchday 35 of the 20/21 season:
league_one_matchday_1_table <- tm_matchday_table(start_year="2020", matchday=1, league_url="https://www.transfermarkt.com/league-one/startseite/wettbewerb/GB3")
```


***

## Team Data

This section will cover off the functions to get team-level data from Transfermarkt.


### Transfer activity by team

To get all the arrivals and departures for a team (or teams) in a season and data regarding the transfer (transfer value, contract length, where they came from/went to, etc), the `tm_team_transfers()` function can be used. This function can return either `summer`, `winter` or `all` for both `transfer_windows`:

```{r team_transfers, eval=FALSE}
# for one team:
bayern <- tm_team_transfers(team_url = "https://www.transfermarkt.com/fc-bayern-munchen/startseite/verein/27/saison_id/2020", transfer_window = "all")

# or for multiple teams:
team_urls <- tm_league_team_urls(country_name = "England", start_year = 2020)
epl_xfers_2020 <- tm_team_transfers(team_url = team_urls, transfer_window = "all")
```


### Squad Player Stats

To get basic statistics (goals, appearances, minutes played, etc) for all games played by players for a squad season, the `tm_squad_stats()` function can be used:

```{r squad_stats, eval=FALSE}
# for one team:
bayern <- tm_squad_stats(team_url = "https://www.transfermarkt.com/fc-bayern-munchen/startseite/verein/27/saison_id/2020")

# or for multiple teams:
team_urls <- tm_league_team_urls(country_name = "England", start_year = 2020)
epl_team_players_2020 <- tm_squad_stats(team_url = team_urls)
```


### Player Valuations

To get player valuations for all teams in a league season, use the `get_player_market_values()` function:

```{r, player_vals, eval=FALSE}
big_5_valuations <- get_player_market_values(country_name = c("England", "Spain", "France", "Italy", "Germany"),
                                       start_year = 2021)
```



***


## Player Data

This section will cover the functions available to aid in the extraction of player data.

### Player Bios

To get information about a player, like their age, foot, where they were born, who they play for, their contract details, social media accounts and a whole lot more, use the `tm_player_bio()` function.:

```{r player_bios, eval=FALSE}
# for a single player 
hazard_bio <- tm_player_bio(player_url = "https://www.transfermarkt.com/eden-hazard/profil/spieler/50202")

# for multiple players:
# can make use of a tm helper function:
burnley_player_urls <- tm_team_player_urls(team_url = "https://www.transfermarkt.com/fc-burnley/startseite/verein/1132/saison_id/2020")
# then pass all those URLs to the tm_player_bio
burnley_bios <- tm_player_bio(player_urls = burnley_player_urls)
```

***


## Club Staff Data

From version 0.4.7, users now have the ability to get historical data for club staff from transfermarkt.

The following two functions can be used, depending on the need (in addition to the helper function `tm_team_staff_urls()` detailed above).

### Club Staff History

You can extract all employees by role in a club's history using `tm_team_staff_history()`.

The list of roles that can be passed to the `staff_roles` argument can be found [here](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/transfermarkt_staff/tm_staff_types.csv)

```{r tm_team_staff_history, eval=FALSE}
# get a list of team URLs for the EPL 2021/22 season
epl_teams <- tm_league_team_urls(country_name = "England", start_year = 2021)

# then use the URLs to pass to the function, and select the role you wish to see results for
club_manager_history <- tm_team_staff_history(team_urls = epl_teams, staff_role = "Manager")
club_caretaker_manager_history <- tm_team_staff_history(team_urls = epl_teams, staff_role = "Caretaker Manager")
```


### Staff Member's History

To be able to get all roles held by a selected staff member(s), the `tm_staff_job_history()` function can be used.

The function accepts one argument, `staff_urls`, which can be extracted using `tm_team_staff_urls()` explained in the helpers section above.

```{r tm_staff_job_history, eval=FALSE}
# get a list of team URLs for the EPL 2021/22 season
epl_teams <- tm_league_team_urls(country_name = "England", start_year = 2021)

# get all EPL goal keeping coaches for the 2021/22 season
epl_gk_coaches <- tm_team_staff_urls(team_urls = epl_teams[1:3], staff_role = "Goalkeeping Coach")
# then you can pass these URLs to the function and get job histories for the selected staff members
epl_gk_coach_job_histories <- tm_staff_job_history(staff_urls = epl_gk_coaches)
```




