---
title: "Extracting data from Understat"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting data from Understat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics from the following popular football (soccer) data sites:

* [FBref](https://fbref.com/en/)
* [Transfermarkt](https://www.transfermarkt.com/)
* [Understat](https://understat.com/)
* [fotmob](https://www.fotmob.com/)


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
```

***

## Usage

Package vignettes have been built to help you get started with the package.

* For functions to extract data from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-fbref-data.html)
* For functions to extract data from Transfermarkt, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-transfermarkt-data.html)
* For functions to extract data from Understat, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-understat-data.html)
* For functions to extract data from fotmob, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-fotmob-data.html)
* For functions to extract data for international matches from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/fbref-data-internationals.html)

This vignette will cover the functions to extract data from understat.com


***

## Understat Helper Functions

There are currently no helper functions for Understat functions.

***

## League Season-Level Data

This section will cover the functions to aid in the extraction of season league statistics from Understat.

The following leagues are currently supported by Understat (these values can be passed in to the `league` arguments of most `understat_` functions):

* "EPL"
* "La liga"
* "Bundesliga"
* "Serie A"
* "Ligue 1"
* "RFPL"


### Match Results

To be able to extract match results from Understat, which not only have results and expected goals, but they also provide a probablity of a team winning.

To extract the data, use the `understat_league_match_results()` function:


```{r understat_match_results, eval=FALSE}
# to get the EPL results:
epl_results <- understat_league_match_results(league = "EPL", season_start_year = 2020)
```


### Season Shooting locations

To get shooting locations for a whole season in supported leagues, use the `understat_league_season_shots()` function:

```{r season_shots, eval=FALSE}
ligue1_shot_location <- understat_league_season_shots(league = "Ligue 1", season_start_year = 2020)
```


***

## Match-Level Data

The following sections outlines the functions available to extract data at the per-match level

### Match Shooting Locations

To get shooting locations for an individual match, use the `understat_match_shots()` function:

```{r match_shots, eval=FALSE}
wba_liv_shots <- understat_match_shots(match_url = "https://understat.com/match/14789")
```


***

## Team Data

This section will cover off the functions to get team-level data from Transfermarkt.


### Team Shooting Locations

To get all shots taken and conceded by a team during a season, use the `understat_team_season_shots()` function:

```{r team_shots, eval=FALSE}
# for one team:
man_city_shots <- understat_team_season_shots(team_url = "https://understat.com/team/Manchester_City/2020")
```

***


## Player Data

This section will cover the functions available to aid in the extraction of player data.

### Player Shooting Locations

To get shooting locations for all games a player has participated in (for as long as Understat has data for), use the `understat_player_shots()` function:

```{r player_bios, eval=FALSE}
raheem_sterling_shots <- understat_player_shots(player_url = "https://understat.com/player/618")
```

***


