---
title: "Extracting data from fotmob"
author: "Tony ElHabr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting data from fotmob}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```

## Overview

This package is designed to allow users to extract various world football results and player statistics from the following popular football (soccer) data sites:

-   [FBref](https://fbref.com/en/)
-   [Transfermarkt](https://www.transfermarkt.com/)
-   [Understat](https://understat.com/)
-   [Fotmob](https://www.fotmob.com/)

## Installation

You can install the CRAN version of [**`worldfootballR`**](https://CRAN.R-project.org/package=worldfootballR) with:

```{r cran-installation, eval=FALSE}
install.packages("worldfootballR")
```

You can install the released version of [**`worldfootballR`**](https://github.com/JaseZiv/worldfootballR/) from [GitHub](https://github.com/JaseZiv/worldfootballR) with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
library(dplyr)
library(tidyr)
```

------------------------------------------------------------------------

## Usage

Package vignettes have been built to help you get started with the package.

-   For functions to extract data from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-fbref-data.html)
-   For functions to extract data from Transfermarkt, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-transfermarkt-data.html)
-   For functions to extract data from Understat, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-understat-data.html)
-   For functions to extract data for international matches from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/fbref-data-internationals.html)
-   For functions to load pre-scraped data, see [here](https://jaseziv.github.io/worldfootballR/articles/load-scraped-data.html)

This vignette will cover the functions to extract data from fotmob.com

------------------------------------------------------------------------

## fotmob Helper Functions

There are currently no helper functions for fotmob functions.

------------------------------------------------------------------------

## League Season-Level Data

fotmob has data for just about every league that you can think of, including all of the Big 5 leagues.

### Team Stats

Use `fotmob_get_season_stats` to retrieve values for a specified stat, season, and league. See the docs for the function for a complete list of `stat_name`s. Note that some stats are only available for either team or player, and that some names are different for teams and players (e.g. `"Expected goals"` for `"team"` and `"Expected goals (xG)"` for `"player"`).

```{r epl_team_xg_2021}
epl_team_xg_2021 <- fotmob_get_season_stats(
  country = "ENG",
  league_name = "Premier League",
  season_name = "2020/2021",
  stat_name = "Expected goals",
  team_or_player = "team"
)

epl_team_xg_2021 %>%
  dplyr::select(
    league_id,
    league_name,
    season_id,
    season_name,
    team_id,
    team_name = participant_name,
    matches_played,
    xg = stat_value,
    g = sub_stat_value
  ) %>%
  dplyr::glimpse()
```

If you know the `league_id`, you can use it instead of passing in values for `country` and `league_name`.

```{r epl_team_xg_2021_w_league_id}
fotmob_get_season_stats(
  league_id = 47,
  season_name = "2020/2021",
  stat_name = "Expected goals",
  team_or_player = "team"
)
```

You can retrieve data for multiple leagues, seasons, or stat types at the same time.

```{r team_xgs_2021}
team_xgs_2021 <- fotmob_get_season_stats(
  country =        c("ITA",     "ESP"),
  league_name =    c("Serie A", "LaLiga"),
  season_name =    c("2020/2021", "2021/2022"),
  stat_name =      c("Expected goals", "xG conceded"),
  team_or_player = "team"
)

## 2 leagues x 20 teams x 2 seasons x 2 stats = 160 rows
team_xgs_2021 %>% nrow()
```

International tournament data can be retrieved, during the right time of year.

```{r cl_team_xg_2021}
## Can only check on CL after group stages and briefly after the final.
m <- lubridate::month(Sys.Date())
if(m >= 1 && m <= 5) {
  fotmob_get_season_stats(
    league_id = 42,
    season_name = "2020/2021",
    stat_name = "Expected goals",
    team_or_player = "team"
  )
}
```

You can potentially get cup data for leagues in the future, although at this time Fotmob does not have this data for most leagues, seasons, or stats.

```{r facup_team_xg_2021, eval=F}
## Not run to avoid error.
fotmob_get_season_stats(
  league_id = 47,
  season_name = "2020/2021",
  stat_name = "Expected goals",
  team_or_player = "team",
  stat_league_name = "FA Cup"
)
```

At this time, only one `team_or_player` can be retrieved at a time.

### Player Stats

You can also use `fotmob_get_season_stats` to retrieve stats for players.

```{r epl_player_xg_2021}
epl_player_xg_2021 <- fotmob_get_season_stats(
  country = "ENG",
  league_name = "Premier League",
  season = "2020/2021",
  stat_name = "Expected goals (xG)",
  team_or_player = "player"
)

epl_player_xg_2021 %>%
  dplyr::select(
    league_id,
    league_name,
    season_id,
    season_name,
    team_id,
    ## NOTE: particiant_id is a typo on behalf of fotmob! We leave it as is.
    player_id = particiant_id,
    player_name = participant_name,
    minutes_played,
    matches_played,
    xg = stat_value,
    g = sub_stat_value
  ) %>%
  dplyr::glimpse()
```

### Match Results

Matches for one or multiple leagues can be retrieved with `fotmob_get_league_matches`. Only the current season's matches can be retrieved with this function.

```{r fotmob_get_league_matches}
league_matches <- fotmob_get_league_matches(
  country =     c("ENG",            "ESP"   ),
  league_name = c("Premier League", "LaLiga")
)

league_matches_unnested <- league_matches %>%
  dplyr::select(match_id = id, home, away) %>%
  tidyr::unnest_wider(c(home, away), names_sep = "_")
dplyr::glimpse(league_matches_unnested)
```

fotmob also lists data for all leagues by date. Use `fotmob_get_matches_by_date` to select matches occurring on specific day(s) and filter down to the league(s) that you care about. Since `fotmob_get_league_matches` does not allow for selecting seasons, the only way to get matches from prior seasons is to use `fotmob_get_matches_by_date`

```{r fotmob_get_matches_by_date}
results <- fotmob_get_matches_by_date(date = c("20210925", "20210926"))
dplyr::glimpse(results)
```

You should see international tournaments in the output from `fotmob_get_matches_by_date()`.

```{r fotmob_get_matches_by_date_int}
results <- fotmob_get_matches_by_date("20220412")
results %>%
  dplyr::filter(name == "Champions League Final Stage", ccode == "INT")
```

### Standings

Standings for one or multiple leagues can be retrieved with `fotmob_get_league_tables`. Currently only the current season's standings can be retrieved.

```{r fotmob_get_league_tables}
league_tables <- fotmob_get_league_tables(
  country =     c("ENG",            "ESP"   ),
  league_name = c("Premier League", "LaLiga")
)
# or
# league_tables <- fotmob_get_league_tables(league_id = c(47, 87))

away_league_tables <- league_tables %>%
  dplyr::filter(table_type == "away")
dplyr::glimpse(away_league_tables)
```

Note that the output of `fotmob_get_league_tables` for a tournament has more columns.

```{r fotmob_get_league_tables_int}
m <- lubridate::month(Sys.Date())
if(m >= 1 && m <= 5) {
  cl_table <- fotmob_get_league_tables(league_id = 42)
  
  cl_table %>%
    dplyr::filter(table_type == "all") %>% 
    dplyr::glimpse()
}
```

------------------------------------------------------------------------

## Match-Level Data

### Match Shooting Locations

To get shooting locations for an individual match along with expected goals (xG), expected goals on target (xGoT), etc., use the `fotmob_get_match_details()` function.

```{r fotmob_match_details}
fotmob_matches <- c(3609994, 3610132)
match_details <- fotmob_get_match_details(fotmob_matches)
dplyr::glimpse(match_details)
```

Be very careful with interpreting `x` and `y` from `on_goal_shot`. They're not on the same scale as `x` and `y` in the top-level of the result returned from `fotmob_get_match_details`!

### Players

You can also extract players from matches with the `fotmob_get_match_players()` function.

```{r fotmob_players}
players <- fotmob_get_match_players(fotmob_matches)
dplyr::glimpse(players)
```

```{r salah_shotmap}
salah <- players %>% dplyr::filter(id == "292462")
salah_shotmap <- salah %>% 
  dplyr::select(player_id = id, shotmap) %>% 
  tidyr::unnest(shotmap)
dplyr::glimpse(salah_shotmap)
```
