---
title: "Extracting fbref URLs"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting fbref URLs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics data from fbref.com and valuations and transfer data from transfermarkt.com.


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
library(dplyr)
```

***


## Usage

The functions available in this package are designed to enable the extraction of world football data.

There are four main categories of data extract functions in this package:

* fbref helper functions to extract key URLs
* Match-level statistics (team and player)
* Season-level statistics (team and player)
* Player market vluations and transfer histories

This vignette will cover the helper functions to extract important URLs from fbref.


### League URLs

To extract the URL of any country's league(s) (provided fbref have data for the league), use the `fb_league_urls()` function.

This function also accepts a `tier` argument. for first-tier leagues, select '1st', for second-tier select '2nd' and so on.

A fill list of countries available can be found in the `worldfootballR_data` repository and can be found [here](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/all_leages_and_cups/all_competitions.csv).

```{r get_league_urls, eval=FALSE}
fb_league_urls(country = "ENG", gender = "M", season_end_year = 2021, tier = '2nd')
```


### Team URLs

To get a list of URLs for each team in a particular season, the `fb_teams_urls()` function can be used:

```{r get_team_urls, eval=FALSE}
fb_teams_urls("https://fbref.com/en/comps/9/Premier-League-Stats")
```



### Player URLs

To get a list of player URLs for a particular team, the `fb_player_urls()` function can be used. The reuslts of this output can be passed through to the player season stat functions `fb_player_season_stats()` and `fb_player_scouting_report()`.

```{r, eval=FALSE}
fb_player_urls("https://fbref.com/en/squads/fd962109/Fulham-Stats")
```

