---
title: "Extracting season data"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting season data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics data from fbref.com and valuations and transfer data from transfermarkt.com.


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
library(dplyr)
```

***


## Usage

The functions available in this package are designed to enable the extraction of world football data.

There are four main categories of data extract functions in this package:

* fbref helper functions to extract key URLs
* Match-level statistics (team and player)
* Season-level statistics (team and player)
* Player market vluations and transfer histories

This vignette will cover season-level data, typically totals for all games played within seasons.

***

## Team Stats

This section will cover the functions to aid in the extraction of season team statistics.

### Get Season Team Stats

The `get_season_team_stats` function allows the user to return a data frame of different stat types for all teams in Domestic leagues [here](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/all_leages_and_cups/all_competitions.csv).

Note, some stats may not be available for all leagues. The big five European leagues should have all of these stats.

The following stat types can be selected:

* *league_table*
* *league_table_home_away*
* *standard*
* *keeper*
* *keeper_adv*
* *shooting*
* *passing*
* *passing_types*
* *goal_shot_creation*
* *defense*
* *possession*
* *playing_time*
* *misc*

```{r get_season_team_stats, eval=FALSE}
# function to extract season teams stats
prem_2020_shooting <- get_season_team_stats(country = "ENG", gender = "M", season_end_year = "2020", tier = "1st", stat_type = "shooting")
# to get shooting stats for the English Championship:
championship_2020_shooting <- get_season_team_stats(country = "ENG", gender = "M", season_end_year = "2020", tier = "2nd", stat_type = "shooting")
```


**More than one league season**

The `get_season_team_stats` function can be used to get data for multiple seasons/leages/genders/etc.

Important to note, this function can only be used for one `stat-type` at a time, however all other parameters can have multiple values:

```{r get_season_team_stats_multiple, eval=FALSE}
big_5_2020_possessions <- get_season_team_stats(country = c("ENG", "ESP", "ITA", "GER", "FRA"),
                                        gender = "M", season_end_year = 2020, tier = "1st", stat_type = "possession")
```


### League Table by Matchdays

To be able to extract league tables for select matchday(s), the below function can be used. 

The function can accept either the country name, season start year and matchday number(s), or for leagues not contained in the [worldfootballR_data](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/transfermarkt_leagues/main_comp_seasons.csv) repository, it can accept the league URL, season start year and matchday number(s).

Note, the league tables are scraped from transfermarkt.

```{r matchday_league_table, eval=FALSE}
# to get the EPL table after matchday 1 of the 20/21 season:
epl_matchday_1_table <- tm_matchday_table(country_name="England", start_year="2020", matchday=1)

# to get the EPL table after each matchdays from matchday 1 to matchday 35 of the 20/21 season:
epl_matchday_1to35_table <- tm_matchday_table(country_name="England", start_year="2020", matchday=c(1:35))

# to get the EPL table after each matchdays from matchday 1 to matchday 35 of the 20/21 season:
league_one_matchday_1_table <- tm_matchday_table(start_year="2020", matchday=1, league_url="https://www.transfermarkt.com/league-one/startseite/wettbewerb/GB3")
```


***

## Player Stats

This section will cover the functions available to aid in the extraction of player season data

### Get Player Scouting Report

The `fb_player_scouting_report()` function takes in a player_url as the input, and returns scouting reports for the player selected. The scouting reports are a new feature on fbref, thus only available for the big five Euro leagues:

```{r get_player_scouting_report}
fb_player_scouting_report("https://fbref.com/en/players/3bb7b8b4/Ederson")
```



### Get Player Season Stats

The `fb_player_season_stats()` function allows for the extraction of season totals for selected player URLs and stat_type.

The stat_types available for use in this function are below:

* *standard*
* *shooting*
* *passing* 
* *passing_types*
* *gca*
* *defense*
* *possession*
* *playing_time*
* *misc*
* *keeper*
* *keeper_adv*

```{r player_season_stats, eval=FALSE}
mo_shooting <- fb_player_season_stats("https://fbref.com/en/players/e342ad68/Mohamed-Salah", stat_type = 'shooting')
```


## The Big 5 Euro Leagues

The `fb_big5_advanced_season_stats()` function allows users to extract data for any of the below listed stat types for either all players, or all teams of the big five European leagues (EPL, La Liga, Ligue 1, Serie A, Bundesliga).

The stat types available for this function are below:

* *standard*
* *shooting*
* *passing*
* *passing_types*
* *gca*
* *defense*
* *possession*
* *playing_time*
* *misc*
* *keepers*
* *keepers_adv*

The function also accepts a season or seasons and whether you want data for the player, or team.

Note that when selecting `team_or_player="team"`, results will be returned for both the team's for and against stats. To filter on this, use the `Team_or_Opponent` column in the resulting data frame, selecting 'team' if you want the team's for stats, or 'opponent' if you want the team's against stats. 


```{r big5_player, eval=FALSE}
big5_player_possession <- fb_big5_advanced_season_stats(season_end_year= 2021, stat_type= "possession", team_or_player= "player")
```

```{r big5_team, eval=FALSE}
big5_team_shooting <- fb_big5_advanced_season_stats(season_end_year= c(2019:2021), stat_type= "shooting", team_or_player= "team")
```


