---
title: "Extracting whole league match data"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting whole league match data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics data from fbref.com and valuations and transfer data from transfermarkt.com.


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
library(dplyr)
```

***


## Usage

The functions available in this package are designed to enable the extraction of world football data.

There are five main categories of data extract functions in this package:

* fbref helper functions to extract key URLs
* Match-level statistics (team and player) for whole leagues
* Season-level statistics (team and player) for whole leagues
* Statistics for individual players and or teams
* Player market vluations and transfer histories

This vignette will cover match-level data for whole leagues. It is broken down into two main sections;

1. Functions to get **team** stats for matches played
3. Functions to get **team OR player** stats for matches played

***

## Match-level data - TEAMS

This section contains examples of functions to use when you want to return statistics for teams.

### Get match urls

To get the match URLs needed to pass in to some of the match-level functions below, `get_match_urls()` can be used:

```{r match_urls, eval=FALSE}
epl_2021_urls <- get_match_urls(country = "ENG", gender = "M", season_end_year = 2021, tier="1st")
```


### Get match results

To get the match results (and additional metadata) for all leagues and comps listed [here](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/all_leages_and_cups/all_competitions.csv), the following function can be used:

```{r get_match_results, eval=FALSE}
# function to extract Serie A match results data
serieA_2020 <- get_match_results(country = "ITA", gender = "M", season_end_year = 2020, tier = "1st")
```

The function can also be used to return match URLs for a non-domestic league season. To use this functionality, simply leave `country = ''` and pass the non-domestic league URL, which can be found at https://fbref.com/en/comps/

```{r get_int_match_results, eval=FALSE}
# for international friendlies:
get_match_results(country = "", gender = "M", season_end_year = 2018, tier = "", non_dom_league_url = "https://fbref.com/en/comps/218/history/Friendlies-M-Seasons")
```

**More than one league season**

The `get_match_results()` function can be used to get data for multiple seasons/leages/genders/etc also:

```{r get_match_results_multiple, eval=FALSE}
big_5_2020_results <- get_match_results(country = c("ENG", "ESP", "ITA", "GER", "FRA"),
                                        gender = "M", season_end_year = 2020, tier = "1st")
```


### Get match report

This function will return similar results to that of `get_match_results()`, however `get_match_report()` will provide some additional information. It will also only provide it for a single match, not the whole season:  

```{r get_match_report, eval=FALSE}
# function to extract match report data
liv_mci_2020 <- get_match_report(match_url = "https://fbref.com/en/matches/47880eb7/Liverpool-Manchester-City-November-10-2019-Premier-League")
```


### Get match summaries

This function will return the main events that occur during a match, including goals, substitutions and red/yellow cards:  

```{r get_match_summary, eval=FALSE}
# function to extract match summary data
liv_mci_2020_summary <- get_match_summary(match_url = "https://fbref.com/en/matches/47880eb7/Liverpool-Manchester-City-November-10-2019-Premier-League")
```


### Get match lineups

This function will return a dataframe of all players listed for that match, including whether they started on the pitch, or on the bench.

From version 0.2.7, this function now also returns some summary performance data for each player that played, including their position, minutes played, goals, cards, etc.

```{r get_match_lineups, eval=FALSE}
# function to extract match lineups
liv_mci_2020_lineups <- get_match_lineups(match_url = "https://fbref.com/en/matches/47880eb7/Liverpool-Manchester-City-November-10-2019-Premier-League")
```



### Get shooting and shot creation events

The below function allows users to extract shooting and shot creation event data for a match or selected matches. The data returned includes who took the shot, when, with which body part and from how far away. Additionally, the player creating the chance and also the creation before this are included in the data.

```{r get_shot_and_creation, eval=FALSE}
shot_one_match <- get_match_shooting(match_url = "https://fbref.com/en/matches/a3eb7a37/Sheffield-United-Wolverhampton-Wanderers-September-14-2020-Premier-League")

test_urls_multiple <- c("https://fbref.com/en/matches/c0996cac/Bordeaux-Nantes-August-21-2020-Ligue-1",
                        "https://fbref.com/en/matches/9cbccb37/Dijon-Angers-August-22-2020-Ligue-1",
                        "https://fbref.com/en/matches/f96cd5a0/Lorient-Strasbourg-August-23-2020-Ligue-1")
shot_multiple_matches <- get_match_shooting(test_urls_multiple)
```


***

## Match-level data - TEAMS OR PLAYERS

This section contains examples of functions to use when you want to return statistics for either teams or players.

### Get advanced match statistics

The `get_advanced_match_stats()` function allows the user to return a data frame of different stat types for matches played.

Note, some stats may not be available for all leagues. The big five European leagues should have all of these stats.

The following stat types can be selected:

* *summary*
* *passing*
* *passing_types*
* *defense*
* *possession*
* *misc*
* *keeper*

The function can be used for either all players individually:

```{r get_advanced_match_stats, eval=FALSE}
test_urls_multiple <- c("https://fbref.com/en/matches/c0996cac/Bordeaux-Nantes-August-21-2020-Ligue-1",
                        "https://fbref.com/en/matches/9cbccb37/Dijon-Angers-August-22-2020-Ligue-1",
                        "https://fbref.com/en/matches/f96cd5a0/Lorient-Strasbourg-August-23-2020-Ligue-1")

advanced_match_stats <- get_advanced_match_stats(match_url = test_urls_multiple, stat_type = "possession", team_or_player = "player")
```


Or used for the team totals for each match:

```{r get_advanced_match_stats_team, eval=FALSE}
test_urls_multiple <- c("https://fbref.com/en/matches/c0996cac/Bordeaux-Nantes-August-21-2020-Ligue-1",
                        "https://fbref.com/en/matches/9cbccb37/Dijon-Angers-August-22-2020-Ligue-1",
                        "https://fbref.com/en/matches/f96cd5a0/Lorient-Strasbourg-August-23-2020-Ligue-1")

advanced_match_stats_team <- get_advanced_match_stats(match_url = test_urls_multiple, stat_type = "passing_types", team_or_player = "team")
```

