---
title: "Extracting data from FBref"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting data from FBref}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics from the following popular football (soccer) data sites:

* [FBref](https://fbref.com/en/)
* [Transfermarkt](https://www.transfermarkt.com/)
* [Understat](https://understat.com/)


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
```

***

## Usage

Package vignettes have been built to help you get started with the package.

* For functions to extract data from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-fbref-data.html)
* For functions to extract data from Transfermarkt, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-transfermarkt-data.html)
* For functions to extract data from Understat, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-understat-data.html)
* For functions to extract data for international matches from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/fbref-data-internationals.html)

This vignette will cover the functions to extract data from FBref.com.


***

## FBref Helper Functions

The following section will outline the various functions available to find different URLs to be able to pass through the FBref suite of functions outlined in this vignette.

### League URLs

To extract the URL of any country's league(s) (provided fbref have data for the league), use the `fb_league_urls()` function.

This function also accepts a `tier` argument. for first-tier leagues, select '1st', for second-tier select '2nd' and so on.

A fill list of countries available can be found in the `worldfootballR_data` repository and can be found [here](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/all_leages_and_cups/all_competitions.csv).

```{r get_league_urls, eval=FALSE}
fb_league_urls(country = "ENG", gender = "M", season_end_year = 2021, tier = '2nd')
```


### Team URLs

To get a list of URLs for each team in a particular season, the `fb_teams_urls()` function can be used:

```{r get_team_urls, eval=FALSE}
fb_teams_urls("https://fbref.com/en/comps/9/Premier-League-Stats")
```


### Player URLs

To get a list of player URLs for a particular team, the `fb_player_urls()` function can be used. The reuslts of this output can be passed through to the player season stat functions `fb_player_season_stats()` and `fb_player_scouting_report()`.

```{r, eval=FALSE}
fb_player_urls("https://fbref.com/en/squads/fd962109/Fulham-Stats")
```


### Get match urls

To get the match URLs needed to pass in to some of the match-level functions below, `get_match_urls()` can be used:

```{r match_urls, eval=FALSE}
epl_2021_urls <- get_match_urls(country = "ENG", gender = "M", season_end_year = 2021, tier="1st")
```



## League Season-Level Data

This section will cover the functions to aid in the extraction of season team statistics.

### Get Season Team Stats

The `get_season_team_stats` function allows the user to return a data frame of different stat types for all teams in Domestic leagues [here](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/all_leages_and_cups/all_competitions.csv).

Note, some stats may not be available for all leagues. The big five European leagues should have all of these stats.

The following stat types can be selected:

* *league_table*
* *league_table_home_away*
* *standard*
* *keeper*
* *keeper_adv*
* *shooting*
* *passing*
* *passing_types*
* *goal_shot_creation*
* *defense*
* *possession*
* *playing_time*
* *misc*

```{r get_season_team_stats, eval=FALSE}
# function to extract season teams stats
prem_2020_shooting <- get_season_team_stats(country = "ENG", gender = "M", season_end_year = "2020", tier = "1st", stat_type = "shooting")
# to get shooting stats for the English Championship:
championship_2020_shooting <- get_season_team_stats(country = "ENG", gender = "M", season_end_year = "2020", tier = "2nd", stat_type = "shooting")
```


#### More than one league season

The `get_season_team_stats` function can be used to get data for multiple seasons/leages/genders/etc.

Important to note, this function can only be used for one `stat-type` at a time, however all other parameters can have multiple values:

```{r get_season_team_stats_multiple, eval=FALSE}
big_5_2020_possessions <- get_season_team_stats(country = c("ENG", "ESP", "ITA", "GER", "FRA"),
                                        gender = "M", season_end_year = 2020, tier = "1st", stat_type = "possession")
```


#### The Big 5 Euro Leagues

The `fb_big5_advanced_season_stats()` function allows users to extract data for any of the below listed stat types for all teams of the big five European leagues (EPL, La Liga, Ligue 1, Serie A, Bundesliga).

The stat types available for this function are below:

* *standard*
* *shooting*
* *passing*
* *passing_types*
* *gca*
* *defense*
* *possession*
* *playing_time*
* *misc*
* *keepers*
* *keepers_adv*

The function also accepts a season or seasons and whether you want data for the player, or team.

Note that when selecting `team_or_player="team"`, results will be returned for both the team's for and against stats. To filter on this, use the `Team_or_Opponent` column in the resulting data frame, selecting 'team' if you want the team's for stats, or 'opponent' if you want the team's against stats. 


```{r big5_team, eval=FALSE}
big5_team_shooting <- fb_big5_advanced_season_stats(season_end_year= c(2019:2021), stat_type= "shooting", team_or_player= "team")
```


***

## Match-Level Data

The following sections outlines the functions available to extract data at the per-match level

### Get match results

To get the match results (and additional metadata) for all leagues and comps listed [here](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/all_leages_and_cups/all_competitions.csv), the following function can be used:

```{r get_match_results, eval=FALSE}
# function to extract Serie A match results data
serieA_2020 <- get_match_results(country = "ITA", gender = "M", season_end_year = 2020, tier = "1st")
```

The function can also be used to return match URLs for a non-domestic league season. To use this functionality, simply leave `country = ''` and pass the non-domestic league URL, which can be found at https://fbref.com/en/comps/

```{r get_int_match_results, eval=FALSE}
# for international friendlies:
get_match_results(country = "", gender = "M", season_end_year = 2018, tier = "", non_dom_league_url = "https://fbref.com/en/comps/218/history/Friendlies-M-Seasons")
```

#### More than one league season

The `get_match_results()` function can be used to get data for multiple seasons/leages/genders/etc also:

```{r get_match_results_multiple, eval=FALSE}
big_5_2020_results <- get_match_results(country = c("ENG", "ESP", "ITA", "GER", "FRA"),
                                        gender = "M", season_end_year = 2020, tier = "1st")
```


### Get match report

This function will return similar results to that of `get_match_results()`, however `get_match_report()` will provide some additional information. It will also only provide it for a single match, not the whole season:  

```{r get_match_report, eval=FALSE}
# function to extract match report data
liv_mci_2020 <- get_match_report(match_url = "https://fbref.com/en/matches/47880eb7/Liverpool-Manchester-City-November-10-2019-Premier-League")
```


### Get match summaries

This function will return the main events that occur during a match, including goals, substitutions and red/yellow cards:  

```{r get_match_summary, eval=FALSE}
# function to extract match summary data
liv_mci_2020_summary <- get_match_summary(match_url = "https://fbref.com/en/matches/47880eb7/Liverpool-Manchester-City-November-10-2019-Premier-League")
```


### Get match lineups

This function will return a dataframe of all players listed for that match, including whether they started on the pitch, or on the bench.

From version 0.2.7, this function now also returns some summary performance data for each player that played, including their position, minutes played, goals, cards, etc.

```{r get_match_lineups, eval=FALSE}
# function to extract match lineups
liv_mci_2020_lineups <- get_match_lineups(match_url = "https://fbref.com/en/matches/47880eb7/Liverpool-Manchester-City-November-10-2019-Premier-League")
```



### Get shooting and shot creation events

The below function allows users to extract shooting and shot creation event data for a match or selected matches. The data returned includes who took the shot, when, with which body part and from how far away. Additionally, the player creating the chance and also the creation before this are included in the data.

```{r get_shot_and_creation, eval=FALSE}
shot_one_match <- get_match_shooting(match_url = "https://fbref.com/en/matches/a3eb7a37/Sheffield-United-Wolverhampton-Wanderers-September-14-2020-Premier-League")

test_urls_multiple <- c("https://fbref.com/en/matches/c0996cac/Bordeaux-Nantes-August-21-2020-Ligue-1",
                        "https://fbref.com/en/matches/9cbccb37/Dijon-Angers-August-22-2020-Ligue-1",
                        "https://fbref.com/en/matches/f96cd5a0/Lorient-Strasbourg-August-23-2020-Ligue-1")
shot_multiple_matches <- get_match_shooting(test_urls_multiple)
```



### Get advanced match statistics

The `get_advanced_match_stats()` function allows the user to return a data frame of different stat types for matches played.

Note, some stats may not be available for all leagues. The big five European leagues should have all of these stats.

The following stat types can be selected:

* *summary*
* *passing*
* *passing_types*
* *defense*
* *possession*
* *misc*
* *keeper*

The function can be used for either all players individually:

```{r get_advanced_match_stats, eval=FALSE}
test_urls_multiple <- c("https://fbref.com/en/matches/c0996cac/Bordeaux-Nantes-August-21-2020-Ligue-1",
                        "https://fbref.com/en/matches/9cbccb37/Dijon-Angers-August-22-2020-Ligue-1",
                        "https://fbref.com/en/matches/f96cd5a0/Lorient-Strasbourg-August-23-2020-Ligue-1")

advanced_match_stats <- get_advanced_match_stats(match_url = test_urls_multiple, stat_type = "possession", team_or_player = "player")
```


Or used for the team totals for each match:

```{r get_advanced_match_stats_team, eval=FALSE}
test_urls_multiple <- c("https://fbref.com/en/matches/c0996cac/Bordeaux-Nantes-August-21-2020-Ligue-1",
                        "https://fbref.com/en/matches/9cbccb37/Dijon-Angers-August-22-2020-Ligue-1",
                        "https://fbref.com/en/matches/f96cd5a0/Lorient-Strasbourg-August-23-2020-Ligue-1")

advanced_match_stats_team <- get_advanced_match_stats(match_url = test_urls_multiple, stat_type = "passing_types", team_or_player = "team")
```

***

## Team-Level Data

This section will cover off the functions to get team-level data from FBref.


### Match Results by team
To get all the results a team(s) has competed in for a season, the following function can be used. The resulting data frame output will include all game results, including any cup games played, and will accept either one, or many team URLs.

```{r team_match_logs, eval=FALSE}
# for single teams:
man_city_2021_url <- "https://fbref.com/en/squads/b8fd03ef/Manchester-City-Stats"
man_city_2021_results <- get_team_match_results(man_city_url)
# get all team URLs for a league
epl_2021_team_urls <- fb_teams_urls("https://fbref.com/en/comps/9/Premier-League-Stats")
epl_2021_team_results <- get_team_match_results(team_url = team_urls)
```

***


## Player-Level Data

This section will cover the functions available to aid in the extraction of player season data.

The examples provided below in a lot of cases have the actual url (player or team) passed to them, however the suite of fbref helper functions outlined in [this helpers vignette](https://jaseziv.github.io/worldfootballR/articles/extract-helper-urls.html) could also be used.


### Get Player Scouting Report

The `fb_player_scouting_report()` function takes in two inputs;

* `player_url` - the URL of the player's main page
* `pos_versus`  which can return the player's comparison against players in their "primary", OR "secondary" position

and returns the full scouting report for the player selected.

```{r get_player_scouting_report}
dplyr::glimpse(fb_player_scouting_report(player_url = "https://fbref.com/en/players/d70ce98e/Lionel-Messi", pos_versus = "primary"))
dplyr::glimpse(fb_player_scouting_report(player_url = "https://fbref.com/en/players/d70ce98e/Lionel-Messi", pos_versus = "secondary"))
```


### Get Player Season Stats

The `fb_player_season_stats()` function allows for the extraction of season totals for selected player URLs and stat_type.

The stat_types available for use in this function are below:

* *standard*
* *shooting*
* *passing* 
* *passing_types*
* *gca*
* *defense*
* *possession*
* *playing_time*
* *misc*
* *keeper*
* *keeper_adv*

```{r player_season_stats, eval=FALSE}
mo_shooting <- fb_player_season_stats("https://fbref.com/en/players/e342ad68/Mohamed-Salah", stat_type = 'shooting')

multiple_playing_time <- fb_player_season_stats(player_url = c("https://fbref.com/en/players/d70ce98e/Lionel-Messi",
                                                "https://fbref.com/en/players/dea698d9/Cristiano-Ronaldo"),
                                 stat_type = "playing_time")
```


### The Big 5 Euro League Players

The `fb_big5_advanced_season_stats()` function allows users to extract data for any of the below listed stat types for all players of the big five European leagues (EPL, La Liga, Ligue 1, Serie A, Bundesliga).

The stat types available for this function are below:

* *standard*
* *shooting*
* *passing*
* *passing_types*
* *gca*
* *defense*
* *possession*
* *playing_time*
* *misc*
* *keepers*
* *keepers_adv*

The function also accepts a season or seasons and whether you want data for the player, or team.


```{r big5_player, eval=FALSE}
big5_player_possession <- fb_big5_advanced_season_stats(season_end_year= 2021, stat_type= "possession", team_or_player= "player")
```


### Player Match Logs

The `fb_player_match_logs()` function allows the user to return a data frame of the match logs of different stat types for a player's matches played in a season.

The following stat types can be selected, depending on the player's position (ie a striker probably won't have "keepers" stats):

* *summary*
* *keepers*
* *passing*
* *passing_types*
* *gca*
* *defense*
* *possession*
* *misc*


```{r player_match_logs, eval=FALSE}
ederson_summary <- fb_player_match_logs("https://fbref.com/en/players/3bb7b8b4/Ederson", season_end_year = 2021, stat_type = 'summary')
```


***


