---
title: "Extracting valuation and transfer data"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting valuation and transfer data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics data from fbref.com and valuations and transfer data from transfermarkt.com.


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
library(dplyr)
```

***


## Usage

The functions available in this package are designed to enable the extraction of world football data.

There are four main categories of data extract functions in this package:

* fbref helper functions to extract key URLs
* Match-level statistics (team and player)
* Season-level statistics (team and player)
* Player market vluations and transfer histories

This vignette will cover the functions to extract player market values and transfer histories from transfermarkt.com.

***

### Transfer and Market Valuations

The following functions allow the user to extract transfer data for players, and also market valuations by league season. The data for these functions comes from transfermarkt.com

*NOTE: All valuations returned are in Euros unless otherwise stated*

#### Market Values

The following function allows the user to extract player market values from transfermarkt.com.

To get a list of the countries available, see the following link https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/transfermarkt_leagues/main_comp_seasons.csv

```{r get_market_vals, eval=FALSE}
prem_vals <- get_player_market_values(country_name = "England", start_year = 2020)
```


If the country/countries aren't available in the main data set, the function can also accept a League URL from transfermarkt.com. To get the league URL, use the filtering options towards the top of transfermarkt.com, select the country and league you want to collect data from, head to that page, and copy the URL.

To extract data for League One in England, the below code will return the player market valuations for the 2020/21 season in League One:

```{r get_unique_league_vals, eval=FALSE}
league_one_vals <- get_player_market_values(start_year = 2020, league_url = "https://www.transfermarkt.com/league-one/startseite/wettbewerb/GB3")
```



#### Transfer history

The following function allows the user to get transfer history for a player or group of players, only requiring the player urls from transfermarkt.com:

```{r, eval=FALSE}
urls <- c("https://www.transfermarkt.com/cristiano-ronaldo/profil/spieler/8198",
          "https://www.transfermarkt.com/alisson/profil/spieler/105470")

transfers <- player_transfer_history(urls)

```


***
