---
title: "Extracting data from Understat"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting data from Understat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics from the following popular football (soccer) data sites:

* [FBref](https://fbref.com/en/)
* [Transfermarkt](https://www.transfermarkt.com/)
* [Understat](https://understat.com/)


## Installation

You can install the CRAN version of [**```worldfootballR```** ](https://CRAN.R-project.org/package=worldfootballR) with:

```{r cran-installation, eval=FALSE}
install.packages("worldfootballR")
```

You can install the released version of [**```worldfootballR```**](https://github.com/JaseZiv/worldfootballR/) from [GitHub](https://github.com/JaseZiv/worldfootballR) with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```


```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
```

***

## Usage

Package vignettes have been built to help you get started with the package.

* For functions to extract data from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-fbref-data.html)
* For functions to extract data from Transfermarkt, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-transfermarkt-data.html)
* For functions to extract data for international matches from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/fbref-data-internationals.html)
* For functions to load pre-scraped data, see [here](https://jaseziv.github.io/worldfootballR/articles/load-scraped-data.html)

This vignette will cover the functions to extract data from understat.com


***

## Understat Helper Functions


### Team Names
To get a list of all available teams names team selected league, use the `understat_avalaible_teams()` function.

You can pass the results of the `understat_avalaible_teams()` function execution to the `understat_team_meta()` function.

```{r team_urls, eval=FALSE}
team_urls <- understat_team_meta(team_name = understat_avalaible_teams(league = 'EPL'))
```


### Team URLs

To get a list of all season team URLs for selected teams, use the `understat_team_meta()` function (note, to get team names, it might be advisable to look at Understat.com's spelling of the team names and pass that through to the function):

```{r team_urls, eval=FALSE}
team_urls <- understat_team_meta(team_name = c("Liverpool", "Manchester City"))
```


***

## League Season-Level Data

This section will cover the functions to aid in the extraction of season league statistics from Understat.

The following leagues are currently supported by Understat (these values can be passed in to the `league` arguments of most `understat_` functions):

* "EPL"
* "La liga"
* "Bundesliga"
* "Serie A"
* "Ligue 1"
* "RFPL"


### Match Results

To be able to extract match results from Understat, which not only have results and expected goals, but they also provide a probability of a team winning.

To extract the data, use the `understat_league_match_results()` function:


```{r understat_match_results}
# to get the EPL results:
epl_results <- understat_league_match_results(league = "EPL", season_start_year = 2020)
dplyr::glimpse(epl_results)
```


### Season Shooting locations

To get shooting locations for a whole season in supported leagues, use the `understat_league_season_shots()` function:

```{r season_shots, eval=FALSE}
ligue1_shot_location <- understat_league_season_shots(league = "Ligue 1", season_start_year = 2020)
```


***

## Match-Level Data

The following sections outlines the functions available to extract data at the per-match level

### Match Shooting Locations

To get shooting locations for an individual match, use the `understat_match_shots()` function:

```{r match_shots}
wba_liv_shots <- understat_match_shots(match_url = "https://understat.com/match/14789")
dplyr::glimpse(wba_liv_shots)
```


***

## Team Data

This section will cover off the functions to get team-level data from Transfermarkt.


### Team Shooting Locations

To get all shots taken and conceded by a team during a season, use the `understat_team_season_shots()` function:

```{r team_shots}
# for one team:
man_city_shots <- understat_team_season_shots(team_url = "https://understat.com/team/Manchester_City/2020")
dplyr::glimpse(man_city_shots)
```


### Team Stat Breakdowns

To get a more granular breakdown of team shooting data for whole seasons, the `understat_team_stats_breakdown()` function can be used. This functions returns a breakdown of team shooting data based on the following groupings:

* Situation
* Formation
* Game state
* Timing
* Shot zones
* Attack speed
* Result

```{r team_breakdown}
#----- Can get data for single teams at a time: -----#
team_breakdown <- understat_team_stats_breakdown(team_urls = "https://understat.com/team/Liverpool/2020")
dplyr::glimpse(team_breakdown)

#----- Or for multiple teams: -----#
# team_urls <- c("https://understat.com/team/Liverpool/2020",
#                "https://understat.com/team/Manchester_City/2020")
# team_breakdown <- understat_team_stats_breakdown(team_urls = team_urls)
```


***


## Player Data

This section will cover the functions available to aid in the extraction of player data.

### Player Shooting Locations

To get shooting locations for all games a player has participated in (for as long as Understat has data for), use the `understat_player_shots()` function:

```{r player_shots}
raheem_sterling_shots <- understat_player_shots(player_url = "https://understat.com/player/618")
dplyr::glimpse(raheem_sterling_shots)
```


### Team Player Season Stats

To get stats for all players of selected teams, run the `understat_team_players_stats()` function.

Note: Team URLs cal be extracted using `understat_team_meta()`.

```{r team_players}
team_players <- understat_team_players_stats(team_url = c("https://understat.com/team/Liverpool/2020", "https://understat.com/team/Manchester_City/2020"))
dplyr::glimpse(team_players)
```


***


