---
title: "Extracting match data"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting match data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics data from fbref.com and valuations and transfer data from transfermarkt.com.


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
library(dplyr)
```

***


## Usage

The functions available in this package are designed to enable the extraction of world football data.

There are three main categories of data extract functions in this package:

* Match-level statistics (team and player)
* Season-level statistics (team and player)
* Player market vluations and transfer histories

This vignette will cover match-level data.

***

## Match-level data

### Get match urls

To get the match URLs needed to pass in to some of the match-level functions below, `get_match_urls()` can be used:

```{r match_urls, eval=FALSE}
epl_2021_urls <- get_match_urls(country = "ENG", gender = "M", season_end_year = 2021)
```


### Get match results

To get the match results (and additional metadata) for all games for a tier-1 league season, the following function can be used:

```{r get_match_results}
# function to extract match results data
serieA_2020 <- get_match_results(country = "ITA", gender = "M", season_end_year = 2020)
glimpse(serieA_2020)
```

**More than one league season**

The `get_match_results` function can be used to get data for multiple seasons/leages/genders/etc also:

```{r get_match_results_multiple, eval=FALSE}
big_5_2020_results <- get_match_results(country = c("ENG", "ESP", "ITA", "GER", "FRA"),
                                        gender = "M", season_end_year = 2020)
```


### Get match report

This function will return similar results to that of `get_match_results()`, however `get_match_report()` will provide some additional information. It will also only provide it for a single match, not the whole season:  

```{r get_match_report}
# function to extract match report data
liv_mci_2020 <- get_match_report(match_url = "https://fbref.com/en/matches/47880eb7/Liverpool-Manchester-City-November-10-2019-Premier-League")
glimpse(liv_mci_2020)
```


### Get match summaries

This function will return the main events that occur during a match, including goals, substitutions and red/yellow cards:  

```{r get_match_summary}
# function to extract match summary data
liv_mci_2020_summary <- get_match_summary(match_url = "https://fbref.com/en/matches/47880eb7/Liverpool-Manchester-City-November-10-2019-Premier-League")
glimpse(liv_mci_2020_summary)
```


### Get advanced match statistics

The `get_advanced_match_stats` function allows the user to return a data frame of different stat types for matches played.

Note, some stats may not be available for all leagues. The big five European leagues should have all of these stats.

The following stat types can be selected:

* *summary*
* *passing*
* *passing_types*
* *defense*
* *possession*
* *misc*
* *keeper*

The function can be used for either all players individually:

```{r get_advanced_match_stats}
test_urls_multiple <- c("https://fbref.com/en/matches/c0996cac/Bordeaux-Nantes-August-21-2020-Ligue-1",
                        "https://fbref.com/en/matches/9cbccb37/Dijon-Angers-August-22-2020-Ligue-1",
                        "https://fbref.com/en/matches/f96cd5a0/Lorient-Strasbourg-August-23-2020-Ligue-1")

advanced_match_stats <- get_advanced_match_stats(match_url = test_urls_multiple, stat_type = "possession", team_or_player = "player")
glimpse(advanced_match_stats)
```


Or used for the team totals for each match:

```{r get_advanced_match_stats_team}
test_urls_multiple <- c("https://fbref.com/en/matches/c0996cac/Bordeaux-Nantes-August-21-2020-Ligue-1",
                        "https://fbref.com/en/matches/9cbccb37/Dijon-Angers-August-22-2020-Ligue-1",
                        "https://fbref.com/en/matches/f96cd5a0/Lorient-Strasbourg-August-23-2020-Ligue-1")

advanced_match_stats_team <- get_advanced_match_stats(match_url = test_urls_multiple, stat_type = "passing_types", team_or_player = "team")
glimpse(advanced_match_stats_team)
```


### Get match lineups

This function will return a dataframe of all players listed for that match, including whether they started on the pitch, or on the bench.  

```{r get_match_lineups}
# function to extract match lineups
liv_mci_2020_lineups <- get_match_lineups(match_url = "https://fbref.com/en/matches/47880eb7/Liverpool-Manchester-City-November-10-2019-Premier-League")
glimpse(liv_mci_2020_lineups)
```
