---
title: "Extracting data from FBref for International Matches"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting data from FBref for International Matches}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics from the following popular football (soccer) data sites:

* [FBref](https://fbref.com/en/)
* [Transfermarkt](https://www.transfermarkt.com/)
* [Understat](https://understat.com/)
* [Fotmob](https://www.fotmob.com/)


## Installation

You can install the CRAN version of [**```worldfootballR```** ](https://CRAN.R-project.org/package=worldfootballR) with:

```{r cran-installation, eval=FALSE}
install.packages("worldfootballR")
```

You can install the released version of [**```worldfootballR```**](https://github.com/JaseZiv/worldfootballR/) from [GitHub](https://github.com/JaseZiv/worldfootballR) with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```


```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
```

***

## Usage

Package vignettes have been built to help you get started with the package.

* For functions to extract data from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-fbref-data.html)
* For functions to extract data from Transfermarkt, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-transfermarkt-data.html)
* For functions to extract data from Understat, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-understat-data.html)
* For functions to extract data from fotmob, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-fotmob-data.html)

This vignette will cover the functions to extract data for international matches from FBref.com.

The functions in this document are all documented in the FBref data vignette, however the method for using these functions for international matches differs from those played in domestic leagues.

**NOTE:**

As of version `0.5.2`, all FBref functions now come with a user-defined pause between page loads to address their new rate limiting. See [this](https://www.sports-reference.com/bot-traffic.html) document for more information. 


***

## Important

To get the competition URLs needed for a lot of the functions in this document, refer to the column `comp_url` for the relevant league / competition you're interested in, in this stored data [file](https://github.com/JaseZiv/worldfootballR_data/blob/master/raw-data/all_leages_and_cups/all_competitions.csv).

***

## Helper Function

### Get match urls

To get the match URLs needed to pass in to some of the match-level functions below, `get_match_urls()` can be used.

```{r match_urls_int, eval=FALSE}
wc_2018_urls <- get_match_urls(country = "", gender = "M", season_end_year = 2018, tier = "", non_dom_league_url = "https://fbref.com/en/comps/1/history/World-Cup-Seasons")

friendly_int_2021_urls <- get_match_urls(country = "", gender = "M", season_end_year = 2021, tier = "", non_dom_league_url = "https://fbref.com/en/comps/218/history/Friendlies-M-Seasons")

euro_2021_urls <- get_match_urls(country = "", gender = "M", season_end_year = 2021, tier = "", non_dom_league_url = "https://fbref.com/en/comps/676/history/European-Championship-Seasons")

copa_2019_urls <- get_match_urls(country = "", gender = "M", season_end_year = 2019, tier = "", non_dom_league_url = "https://fbref.com/en/comps/685/history/Copa-America-Seasons")
```


***

## Match-Level Data

The following sections outlines the functions available to extract data at the per-match level

### Get match results

To get the match results (and additional metadata), the following function can be used.

To use this functionality, simply leave `country = ''` and pass the non-domestic league URL

```{r get_int_match_results_int, eval=FALSE}
# euro 2016 results
euro_2016_results <- get_match_results(country = "", gender = "M", season_end_year = 2016, tier = "", non_dom_league_url = "https://fbref.com/en/comps/676/history/European-Championship-Seasons")

# 2019 Copa America results:
copa_2019_results <- get_match_results(country = "", gender = "M", season_end_year = 2019, non_dom_league_url = "https://fbref.com/en/comps/685/history/Copa-America-Seasons")

# for international friendlies:
international_results <- get_match_results(country = "", gender = "M", season_end_year = 2021, tier = "", non_dom_league_url = "https://fbref.com/en/comps/218/history/Friendlies-M-Seasons")
```



### Get match report

This function will return similar results to that of `get_match_results()`, however `get_match_report()` will provide some additional information. It will also only provide it for a single match, not the whole season:  

```{r get_match_report_int, eval=FALSE}
# function to extract match report data for 2018 world cup
wc_2018_report <- get_match_report(match_url = wc_2018_urls)
# function to extract match report data for 2021 international friendlies
friendlies_report <- get_match_report(match_url = friendly_int_2021_urls)
```


### Get match summaries

This function will return the main events that occur during a match, including goals, substitutions and red/yellow cards:  

```{r get_match_summary_int, eval=FALSE}
# first get the URLs for the 2016 Euros
euro_2016_match_urls <- get_match_urls(country = "", gender = "M", season_end_year = 2016, tier = "", non_dom_league_url = "https://fbref.com/en/comps/676/history/European-Championship-Seasons")

# then pass these to the function to get match summaries:
euro_2016_events <- get_match_summary(euro_2016_match_urls)
```


### Get match lineups

This function will return a dataframe of all players listed for that match, including whether they started on the pitch, or on the bench.

From version 0.2.7, this function now also returns some summary performance data for each player that played, including their position, minutes played, goals, cards, etc.

```{r get_match_lineups_int, eval=FALSE}
# function to extract match lineups
copa_2019_lineups <- get_match_lineups(match_url = copa_2019_urls)
```



### Get shooting and shot creation events

The below function allows users to extract shooting and shot creation event data for a match or selected matches. The data returned includes who took the shot, when, with which body part and from how far away. Additionally, the player creating the chance and also the creation before this are included in the data.

```{r get_shot_and_creation_int, eval=FALSE}
shots_wc_2018 <- get_match_shooting(wc_2018_urls)
```



### Get advanced match statistics

The `get_advanced_match_stats()` function allows the user to return a data frame of different stat types for matches played.

Note, some stats may not be available for all comps.

The following stat types can be selected:

* *summary*
* *passing*
* *passing_types*
* *defense*
* *possession*
* *misc*
* *keeper*

The function can be used for either all players individually:

```{r get_advanced_match_stats_int, eval=FALSE}
advanced_match_stats_player <- get_advanced_match_stats(match_url = wc_2018_urls, stat_type = "possession", team_or_player = "player")
```


Or used for the team totals for each match:

```{r get_advanced_match_stats_team_int, eval=FALSE}
advanced_match_stats_team <- get_advanced_match_stats(match_url = wc_2018_urls, stat_type = "passing_types", team_or_player = "team")
```

***


