---
title: "Load Scraped Data Functions"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Load Scraped Data Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics from the following popular football (soccer) data sites:

* [FBref](https://fbref.com/en/)
* [Transfermarkt](https://www.transfermarkt.com/)
* [Understat](https://understat.com/)
* [Fotmob](https://www.fotmob.com/)


## Installation

You can install the CRAN version of [**```worldfootballR```** ](https://CRAN.R-project.org/package=worldfootballR) with:

```{r cran-installation, eval=FALSE}
install.packages("worldfootballR")
```

You can install the released version of [**```worldfootballR```**](https://github.com/JaseZiv/worldfootballR/) from [GitHub](https://github.com/JaseZiv/worldfootballR) with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```


```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
library(dplyr)
```

***

## Usage

Package vignettes have been built to help you get started with the package.

* For functions to extract data from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-fbref-data.html)
* For functions to extract data from Transfermarkt, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-transfermarkt-data.html)
* For functions to extract data from Understat, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-understat-data.html)
* For functions to extract data from fotmob, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-fotmob-data.html)
* For functions to extract data for international matches from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/fbref-data-internationals.html)

This vignette will cover the functions to load scraped data from the [`worldfootballR_data`](https://github.com/JaseZiv/worldfootballR_data) data repository.

**NOTE:**

As of version `0.5.2`, all FBref functions now come with a user-defined pause between page loads to address their new rate limiting. See [this](https://www.sports-reference.com/bot-traffic.html) document for more information. 


***

## Load FBref

The following section demonstrates the different loading functions of FBref data.

### Load FBref match results

To load pre-scraped match results for all years the data is available, the `load_match_results()` function can be used. This data is scheduled to be updated most days and a print statement will inform the user of when the data was last updated. All domestic leagues are included in the data repository. 

This is the load function equivalent of `get_match_results()`.

```{r load_match_results}
eng_match_results <- load_match_results(country = "ENG", gender = c("M", "F"), season_end_year = c(2020:2022), tier = "1st")
dplyr::glimpse(eng_match_results)
```


### Load FBref match results for Cups and International Comps

Similarly, to load pre-scraped match results for cups and international matches in all years the data is available, the `load_match_comp_results()` function can be used. This data is scheduled to be updated most days and a print statement will inform the user of when the data was last updated. 

The following list of competitions (`comp_name`) are available:

```{r comp_list, echo=FALSE}
seasons <- read.csv("https://raw.githubusercontent.com/JaseZiv/worldfootballR_data/master/raw-data/all_leages_and_cups/all_competitions.csv", stringsAsFactors = F)

# the below cups are one off matches so we don't need scores and fixtures for these:
exclusion_cups <- c("UEFA Super Cup", "FA Community Shield", "Supercopa de España", "Trophée des Champions", "DFL-Supercup", "Supercoppa Italiana")

latest_cup_seasons <- seasons %>%
  # filtering out things that aren't domestic leagues:
  filter(!stringr::str_detect(.data$competition_type, "Leagues"),
         # and also the single match type cup games:
         !.data$competition_name %in% exclusion_cups) %>% 
  group_by(competition_name) %>% slice_max(season_end_year) %>% 
  distinct() %>% 
  select(competition_type,competition_name,country,gender,governing_body,first_season,last_season,tier)

latest_cup_seasons %>% pull(competition_name)
```




```{r load_match_comp_results}
cups <- c("FIFA Women's World Cup","FIFA World Cup")
world_cups <- load_match_comp_results(comp_name = cups)
dplyr::glimpse(world_cups)
```


### Load FBref big 5 league advanced season stats

To load pre-scraped advanced stats for the big five European leagues for either teams or players, the `load_fb_big5_advanced_season_stats()` can be used. This data is scheduled to be updated most days and a print statement will inform the user of when the data was last updated.

This is the load function equivalent of `get_match_results()`.

```{r, load_adv_stats}
all_season_player <- load_fb_big5_advanced_season_stats(stat_type = "defense", team_or_player = "player")
current_season_player <- load_fb_big5_advanced_season_stats(season_end_year = 2022, stat_type = "defense", team_or_player = "player")

all_season_team <- load_fb_big5_advanced_season_stats(stat_type = "defense", team_or_player = "team")
current_season_team <- load_fb_big5_advanced_season_stats(season_end_year = 2022, stat_type = "defense", team_or_player = "team")
```



***

## Load Understat

The following section demonstrates the different loading functions of Understat data.

### Load League Shots

To be able to rapidly load pre-collected chooting locations for whole leagues, the `load_understat_league_shots()` functions is now available. Supported leagues on Understat are:

* "EPL"
* "La liga"
* "Bundesliga"
* "Serie A"
* "Ligue 1"
* "RFPL"

This is effectively the loading equivalent of the `understat_league_season_shots()` function, however rather than needing to be scraped a season at a time, this data loads rapidly for all seasons for the selected league since the 2014/15 seasons.

```{r}
serie_a_shot_locations <- load_understat_league_shots(league = "Serie A")
dplyr::glimpse(serie_a_shot_locations)
```

## Load fotmob

The following section demonstrates the different loading functions of Understat data.

### Load fotmob Big 5 Match Shots

`load_fotmob_match_details()` returns match details dating back to the 2020/21 season (some matches may be missing) for each of the following leagues:

-   EURO: 50
-   Champions League: 42
-   Copa America: 44
-   Europa League: 73
-   Premier League: 47
-   1.  Bundesliga: 54
-   LaLiga: 87
-   Ligue 1: 53
-   MLS: 130
-   Serie A: 55

This is effectively the loading equivalent of the `fotmob_get_match_details()` function for all matches for these leagues back to the beginning of the 2020/21 season.

```{r}
epl_match_details <- load_fotmob_match_details(
  country = "ENG",
  league_name = "Premier League"
)
## or
## load_fotmob_match_details(league_id = 47)
dplyr::glimpse(epl_match_details)

## multiple leagues at once
epl_ll_match_details <- load_fotmob_match_details(league_id = c(47, 87))
```

Note that fotmob does not currently have match details prior to 2020/21.

`load_fotmob_matches_by_date()` can be used to retrieve fotmob match ids dating back to August 2017.

```{r}
epl_matches <- load_fotmob_matches_by_date(
  country = "ENG",
  league_name = "Premier League"
)

epl_ll_matches <- load_fotmob_matches_by_date(league_id = c(47, 87))
```



