---
title: "Load Scraped Data Functions"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Load Scraped Data Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics from the following popular football (soccer) data sites:

* [FBref](https://fbref.com/en/)
* [Transfermarkt](https://www.transfermarkt.com/)
* [Understat](https://understat.com/)
* [Fotmob](https://www.fotmob.com/)


## Installation

You can install the CRAN version of [**```worldfootballR```** ](https://CRAN.R-project.org/package=worldfootballR) with:

```{r cran-installation, eval=FALSE}
install.packages("worldfootballR")
```

You can install the released version of [**```worldfootballR```**](https://github.com/JaseZiv/worldfootballR/) from [GitHub](https://github.com/JaseZiv/worldfootballR) with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```


```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
```

***

## Usage

Package vignettes have been built to help you get started with the package.

* For functions to extract data from Transfermarkt, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-transfermarkt-data.html)
* For functions to extract data from Understat, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-understat-data.html)
* For functions to extract data from fotmob, see [here](https://jaseziv.github.io/worldfootballR/articles/extract-fotmob-data.html)
* For functions to extract data for international matches from FBref, see [here](https://jaseziv.github.io/worldfootballR/articles/fbref-data-internationals.html)
* For functions to load pre-scraped data, see [here](https://jaseziv.github.io/worldfootballR/articles/load-scraped-data.html)

This vignette will cover the functions to load scraped data from the [`worldfootballR_data`](https://github.com/JaseZiv/worldfootballR_data) data repository.

**NOTE:**

As of version `0.5.2`, all FBref functions now come with a user-defined pause between page loads to address their new rate limiting. See [this](https://www.sports-reference.com/bot-traffic.html) document for more information. 


***

## Load FBref match results

To load pre-scraped match results for all years the data is available, the `load_match_results()` function can be used. This data is scheduled to be updated most days and a print statement will inform the user of when the data was last updated. All domestic leagues are included in the data repository. 

This is the load function equivalent of `get_match_results()`.

```{r load_match_results}
eng_match_results <- load_match_results(country = "ENG", gender = c("M", "F"), season_end_year = c(2020:2022), tier = "1st")
dplyr::glimpse(eng_match_results)
```


## Load FBref big 5 league advanced seasdon stats

To load pre-scraped advanced stats for the big five European leagues for either teams or players, the `load_fb_big5_advanced_season_stats()` can be used. This data is scheduled to be updated most days and a print statement will inform the user of when the data was last updated.

This is the load function equivalent of `get_match_results()`.

```{r, load_adv_stats}
all_season_player <- load_fb_big5_advanced_season_stats(stat_type = "defense", team_or_player = "player")
current_season_player <- load_fb_big5_advanced_season_stats(season_end_year = 2022, stat_type = "defense", team_or_player = "player")

all_season_team <- load_fb_big5_advanced_season_stats(stat_type = "defense", team_or_player = "team")
current_season_team <- load_fb_big5_advanced_season_stats(season_end_year = 2022, stat_type = "defense", team_or_player = "team")
```


