---
title: "Extracting data for individual players or teams"
author: "Jason Zivkovic"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting data for individual players or teams}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message=FALSE,
  warning=FALSE
)
```


## Overview

This package is designed to allow users to extract various world football results and player statistics data from fbref.com and valuations and transfer data from transfermarkt.com.


## Installation

You can install the `worldfootballR` package from github with:

```{r gh-installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("JaseZiv/worldfootballR")
```

```{r load_libs, warning=FALSE, message=FALSE}
library(worldfootballR)
library(dplyr)
```

***


## Usage

The functions available in this package are designed to enable the extraction of world football data.

There are five main categories of data extract functions in this package:

* fbref helper functions to extract key URLs
* Match-level statistics (team and player) for whole leagues
* Season-level statistics (team and player) for whole leagues
* Statistics for individual players and or teams
* Player market vluations and transfer histories

This vignette will cover the functions to extract data for individual players and or teams.


***

## Player Stats

This section will cover the functions available to aid in the extraction of player season data.

The examples provided below in a lot of cases have the actual url (player or team) passed to them, however the suite of fbref helper functions outlined in [this helpers vignette](https://jaseziv.github.io/worldfootballR/articles/extract-helper-urls.html) could also be used.

### Get Player Scouting Report

The `fb_player_scouting_report()` function takes in a player_url as the input, and returns scouting reports for the player selected. The scouting reports are a new feature on fbref, thus only available for the big five Euro leagues:

```{r get_player_scouting_report}
fb_player_scouting_report("https://fbref.com/en/players/3bb7b8b4/Ederson")
```



### Get Player Season Stats

The `fb_player_season_stats()` function allows for the extraction of season totals for selected player URLs and stat_type.

The stat_types available for use in this function are below:

* *standard*
* *shooting*
* *passing* 
* *passing_types*
* *gca*
* *defense*
* *possession*
* *playing_time*
* *misc*
* *keeper*
* *keeper_adv*

```{r player_season_stats, eval=FALSE}
mo_shooting <- fb_player_season_stats("https://fbref.com/en/players/e342ad68/Mohamed-Salah", stat_type = 'shooting')
```


### Player Match Logs

The `fb_player_match_logs()` function allows the user to return a data frame of the match logs of different stat types for a player's matches played in a season.

The following stat types can be selected, depending on the player's position (ie a striker probably won't have "keepers" stats):

* *summary*
* *keepers*
* *passing*
* *passing_types*
* *gca*
* *defense*
* *possession*
* *misc*


```{r player_match_logs, eval=FALSE}
ederson_summary <- fb_player_match_logs("https://fbref.com/en/players/3bb7b8b4/Ederson", season_end_year = 2021, stat_type = 'summary')
```


***


## Team Stats

To get all the results a team(s) has competed in for a season, the following function can be used. The resulting dataframe output will include all game results, including any cup games played, and will accept either one, or many team URLs.

```{r team_match_logs, eval=FALSE}
# for single teams:
man_city_2021_url <- "https://fbref.com/en/squads/b8fd03ef/Manchester-City-Stats"
man_city_2021_results <- get_team_match_results(man_city_url)
# get all team URLs for a league
epl_2021_team_urls <- fb_teams_urls("https://fbref.com/en/comps/9/Premier-League-Stats")
epl_2021_team_results <- get_team_match_results(team_url = team_urls)
```


***
